$R^*$: A robust MCMC convergence diagnostic with uncertainty using gradient-boosted machines

### 2020-09-01 write_up/arxiv_submission_aug_2020/ms_arxiv_2020.tex
arxiv submission done; BA submission underway

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/unmixed_1.pdf}}
	\caption{\textbf{Chain prediction based on the marginal distribution of a single parameter.} A shows the path of two chains that have mixed (with marginal distribution to the right of panel); B shows two chains that have not mixed.}
	\label{fig:marginal}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/unmixed_2.pdf}}
	\caption{\textbf{Chain prediction based on the joint distribution of two parameters where each chain's marginals are the same.} A shows the path of two chains that have mixed resulting in similar sampling distributions (to the right and above each panel); B shows two chains that have not mixed.}
	\label{fig:joint}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/ar1.pdf}}
	\caption{\textbf{Autoregressive example.} A shows how the GBM's classifications vary according to the draw's value for an example model fit; B shows $R^*$ values generated by Algorithm \ref{alg:R_star} across 1000 replicate datasets; C shows corresponding rank-normalised split-$\widehat{R}$ values as calculated in \cite{vehtari2019rank} for each of the 1000 replicates; and D shows 1000 $R^*$ samples as generated by Algorithm \ref{alg:R_star_uncertainty} for two series: the ``unmixed'' dataset being the same as used for figures A-C; the ``mixed'' where all chains have the same distribution as described in \S\ref{sec:heterogeneity}. Note that, in D, only a single series of each series type is used to generate distribution.}
	\label{fig:ar1}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=0.6\textwidth]{../output/ar1_samplesize.pdf}}
	\caption{\textbf{Autoregressive example: $R^*$ stochasticity.} The horizontal axis shows sample size; the vertical axis shows the value of $R^*$ calculated as per Algorithm \ref{alg:R_star} applied to chains split into two halves. Grey points show the value of $R^*$ for each replicate (jitter was added to point positions). Black points show the median $R^*$ value; upper and lower whiskers show 2.5\% and 97.5\% quantiles. For each sample size, a single dataset was created and used for all $R^*$ calculations.}
	\label{fig:ar1_samplesize}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=0.8\textwidth]{../output/ar1_numchains.pdf}}
	\caption{\textbf{Autoregressive example: sensitivity to number of chains.} The horizontal axis shows the number of chains used in the data generating process described in \S\ref{sec:heterogeneity_numchains}. The vertical axis shows the value of $R^*$ as calculated by Algorithm \ref{alg:R_star} (left panel) on chains split into two halves, and rank-normalised split-$\widehat{R}$ as calculated in \cite{vehtari2019rank} (right panel). Grey points indicate the values of both convergence measures calculated for each replicate; horizontal jitter has been added to points. The point-ranges shown indicate the 25\%, 50\% and 75\% quantiles across 50 replicates at each number of chains.}
	\label{fig:ar1_numchains}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/ar1_sensitivity.pdf}}
	\caption{\textbf{Autoregressive example: sensitivity to GBM hyperparameters.} The rows correspond to GBM models fit using different hyperparameter sets (as indicated for interaction depth and number of trees and using a shrinkage of 0.1 and a minimum observations per node of 10 in all cases). The horizontal axis shows the $R^*$ values generated using Algorithm \ref{alg:R_star} using split chains; points have jitter added and are coloured according to whether they exceed the threshold $R^*=1$, with the percentage below this threshold annotated. The point-ranges shown indicate the 25\%, 50\% and 75\% quantiles for each model.}
	\label{fig:ar1_sensitivity}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=0.6\textwidth]{../output/bivariate.pdf}}
	\caption{\textbf{Bivariate normal example.} The distribution for $R^*$ across 1000 draws.}
	\label{fig:bivariate}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1\textwidth]{../output/mvt_three.pdf}}
	\caption{\textbf{Multivariate normal example with 250 dimensions.} A shows $R^*$ distributions obtained for two MCMC samples (of differing numbers of draws: 400 and 10,000) from the centered parameterisation (``cp'') and one from the non-centered version (``ncp''; with 10,000 draws); B shows the rank-normalised split-$\widehat{R}$ values for all parameters from the same MCMC runs as in A; C shows variable importance versus $\widehat{R}$ for each parameter; and D shows variable importance versus bulk- and tail-ESS as calculated by \cite{vehtari2019rank}. In A, 1000 $R^*$ draws by Algorithm \ref{alg:R_star_uncertainty} are shown for each MCMC run. In plots C and D, horizontal jitter was added to the points and a loess fit line with standard errors overlaid.}
	\label{fig:mvt}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/cauchy.pdf}}
	\caption{\textbf{Cauchy example.} Rows show convergence results for MCMC runs with 1000 (top) and 10,000 (bottom; obtained by thinning iterations by a factor of 3) post-warm-up iterations (each with half iterations discarded as warm-up) for each of 4 chains. Columns show the $R^*$ distributions (left) and rank-normalised split-$\widehat{R}$ values across all parameters (right). Shadings indicate different model paramerisations as indicated in legend.}
	\label{fig:cauchy}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/cauchy_convergence.pdf}}
	\caption{\textbf{Measuring convergence for the Cauchy model.} A shows a measure of convergence, the mean quantile $R^2$, that requires knowing the target distribution; B shows the maximum value of split-$\widehat{R}$ across each of the 50 dimensions of the target; C shows the multivariate split-$\widehat{R}$ value; D shows the value of split-$R^*$ as calculated by Algorithm \ref{alg:R_star}; and E and F show tail- and bulk-ESS. Dashed lines indicate recommended thresholds for each convergence statistic.}
	\label{fig:cauchy_convergence}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/eight_schools.pdf}}
	\caption{\textbf{Eight schools example: $R^*$ distributions.} A shows draws from the $R^*$ distribution when splitting chains in two (resulting in 8 chains); B shows the same but using the 4 original chains; C shows rank-normalised $\widehat{R}$ for original 4 chains versus those for the 8 chains case for all ten parameters defined by the centered model -- in this case, we plot horizontal and vertical dashed lines to illustrate the $\widehat{R}=1.01$ cutoff and a $y=x$ line. The legend inset in panel C provides a key for all panels. The MCMC samples comprised 2000 draws in all cases, with 1000 used as post-warm-up iterations. In panels A and B, the plots show 1000 $R^*$ draws using Algorithm \ref{alg:R_star_uncertainty} for each parameterisation.}
	\label{fig:eight_schools}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/eight_schools_r_star_quantiles.pdf}}
	\caption{\textbf{Eight schools example: quantile $R^*$ plots.} Column A shows plots for $\mu$; Column B for $\tau$. In each column, we show the path of the four individual chains above and 40 $R^*$ draws obtained using Algorithm \ref{alg:R_star_uncertainty} for each parameter quantile below. Above the top row, we show the marginal distribution of each chain estimated via kernel density estimation using Gaussian kernels. Note that, in the bottom plots, jitter has been added to the data points.}
	\label{fig:eight_schools_r_star_quantiles}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1\textwidth]{../output/mvt_wide_both.pdf}}
	\caption{\textbf{Wide data examples.} A shows the $R^*$ distribution for the 250-dimensional example in \S\ref{sec:wide} with 250 post-warm-up iterations per chain from Stan's NUTS algorithm across both model parameterisations; B shows $R^*$ distribution for the 10,000-dimensional example with 400 and 1000 MCMC iterations per chain (although the first half of these were discarded as warm-up). In all cases, 1000 draws of $R^*$ are plotted as generated by Algorithm \ref{alg:R_star_uncertainty} for a single MCMC run composed of 4 chains.}
	\label{fig:mvt_wide_both}
\end{figure}

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1\textwidth]{../output/wide_both_diagnostics.pdf}}
	\caption{\textbf{Wide data 250-dimensional example: established diagnostics.} The top row shows the results for the centered parameterisation; the bottom row for the non-centered. Column A shows split-$\widehat{R}$; columns B and C show the bulk- and tail-ESS; in each case the statistics are displayed for all model parameters and were calculated using 250 post-warm-up draws from Stan's NUTS algorithm. Note, that it is possible for the ESS to exceed the actual sample size if there is negative autocorrelation in the Markov chains' values. In both cases, the results correspond to a single MCMC run composed of 4 chains.}
	\label{fig:wide_both_diagnostics}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/trends_all_dim.pdf}}
	\caption{\textbf{Univariate trends example.} Column A represents results for $R^*$ calculated using Algorithm \ref{alg:R_star} on the 4 chains; column B  shows the same calculation after each chain is split into two halves. The rows present the differing sample sizes. The horizontal axis measures (half) the change in mean across the whole sample: so a value ``1'' indicates the mean increases by 2 units from the start to end of sampling. At each parameter set, 10 replicates were run and jitter was added to the points.}
	\label{fig:trends_all_dim}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/trends_one_dim.pdf}}
	\caption{\textbf{Multivariate trends example with split-$R^*$.} The columns present different dimensionalities of the target distribution; the rows present different sample sizes. The horizontal axis measures (half) the change in mean across the single dimension that had a trend added to it: a value ``1'' indicates its mean increases by 2 units from the start to end of sampling; all other dimensions (if dimensions exceeded 1) had stationary distributions. At each parameter set, 10 replicates were run and jitter was added to the points.}
	\label{fig:trends_one_dim}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/trends_joint_distribution.pdf}}
	\caption{\textbf{Bivariate normal with trending correlation example.} Column A shows the results for split-$\widehat{R}$; column B for $R^*$; the rows present the differing sample sizes. The horizontal axis measures (half) the change in correlation across the whole sample: so a value ``0.5'' indicates the correlation increases by 1 unit (from -0.5 to 0.5) from the start to end of sampling. At each parameter set, 10 replicates were run and jitter was added to the points.}
	\label{fig:trends_joint_distribution}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/trends_ar1.pdf}}
	\caption{\textbf{Non-stationary distribution: AR(1) example.} The horizontal axis measures the autocorrelation of the AR(1) processes; the vertical axis shows the value of $R^*$ calculated on chains split in half; each panel shows a different number of iterations. At each parameter set, 10 replicates were run and jitter was added to the points. The black line shows $R^*=1$.}
	\label{fig:trends_ar1}
\end{figure}

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/trends_ar1_transposed.pdf}}
	\caption{\textbf{Non-stationary distribution: AR(1) example alternative view.} The horizontal axis measures the number of iterations; the vertical axis shows the value of $R^*$ calculated on chains split in half; each panel shows a different value of autocorrelation. At each parameter set, 10 replicates were run and jitter was added to the points. The black dots show the median $R^*$ values at each parameter set; upper and lower whiskers show 2.5\% and 97.5\% quantiles. The black horizontal line shows $R^*=1$. }
	\label{fig:trends_ar1_transposed}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/ovarian.pdf}}
	\caption{\textbf{Ovarian example.} In all plots, we show the results from two model runs: one with 4 chains run with 800 thinned post-warm-up iterations (9000 total iterations with 1000 discarded as warm-up; 8000 post-warm-up iterations thinned by a factor of 10); another with 16 chains run with 1000 post-warm-up iterations each (500 warm-up iterations discarded and no thinning). In A, $R^*$ distributions (with 1000 draws each using Algorithm \ref{alg:R_star_uncertainty}) are shown; B shows rank-normalised split-$\widehat{R}$ values across all parameters; C shows bulk-ESS across all parameters; and D shows tail-ESS across parameters.}
	\label{fig:ovarian}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{../output/prostate.pdf}}
	\caption{\textbf{Prostate example.} In all plots, we show the results from two model runs: one with 4 chains run with 800 thinned post-warm-up iterations (9000 total iterations with 1000 discarded as warm-up; 8000 post-warm-up iterations thinned by a factor of 10); another with 16 chains run with 1000 post-warm-up iterations each (500 warm-up iterations discarded and no thinning). In A, $R^*$ distributions (with 1000 draws each using Algorithm \ref{alg:R_star_uncertainty}) are shown -- for the long run, these were calculated after thinning the parameters by a factor of 5; B shows rank-normalised split-$\widehat{R}$ values across all parameters; C shows bulk-ESS across all parameters; and D shows tail-ESS across parameters.}
	\label{fig:prostate}
\end{figure}

>>>>

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{unmixed_1.pdf}}
	\caption{\textbf{Chain prediction based on the marginal distribution of a single parameter.} A shows the path of two chains that have mixed (with marginal distribution to the right of panel); B shows two chains that have not mixed.}
	\label{fig:marginal}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1.0\textwidth]{unmixed_2.pdf}}
	\caption{\textbf{Chain prediction based on the joint distribution of two parameters where each chain's marginals are the same.} A shows the path of two chains that have mixed resulting in similar sampling distributions (to the right and above each panel); B shows two chains that have not mixed.}
	\label{fig:joint}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{ar1.pdf}}
	\caption{\textbf{Autoregressive example.} A shows how the GBM's classifications vary according to the draw's value for an example model fit; B shows $R^*$ values generated by Algorithm \ref{alg:R_star} across 1000 replicate datasets; C shows corresponding rank-normalised split-$\widehat{R}$ values as calculated in \cite{vehtari2019rank} for each of the 1000 replicates; and D shows 1000 $R^*$ samples as generated by Algorithm \ref{alg:R_star_uncertainty} for two series: the ``unmixed'' dataset being the same as used for figures A-C; the ``mixed'' where all chains have the same distribution as described in \S\ref{sec:heterogeneity}. Note that, in D, only a single series of each series type is used to generate distribution.}
	\label{fig:ar1}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=0.6\textwidth]{ar1_samplesize.pdf}}
	\caption{\textbf{Autoregressive example: $R^*$ stochasticity.} The horizontal axis shows sample size; the vertical axis shows the value of $R^*$ calculated as per Algorithm \ref{alg:R_star} applied to chains split into two halves. Grey points show the value of $R^*$ for each replicate (jitter was added to point positions). Black points show the median $R^*$ value; upper and lower whiskers show 2.5\% and 97.5\% quantiles. For each sample size, a single dataset was created and used for all $R^*$ calculations.}
	\label{fig:ar1_samplesize}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=0.8\textwidth]{ar1_numchains.pdf}}
	\caption{\textbf{Autoregressive example: sensitivity to number of chains.} The horizontal axis shows the number of chains used in the data generating process described in \S\ref{sec:heterogeneity_numchains}. The vertical axis shows the value of $R^*$ as calculated by Algorithm \ref{alg:R_star} (left panel) on chains split into two halves, and rank-normalised split-$\widehat{R}$ as calculated in \cite{vehtari2019rank} (right panel). Grey points indicate the values of both convergence measures calculated for each replicate; horizontal jitter has been added to points. The point-ranges shown indicate the 25\%, 50\% and 75\% quantiles across 50 replicates at each number of chains.}
	\label{fig:ar1_numchains}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{ar1_sensitivity.pdf}}
	\caption{\textbf{Autoregressive example: sensitivity to GBM hyperparameters.} The rows correspond to GBM models fit using different hyperparameter sets (as indicated for interaction depth and number of trees and using a shrinkage of 0.1 and a minimum observations per node of 10 in all cases). The horizontal axis shows the $R^*$ values generated using Algorithm \ref{alg:R_star} using split chains; points have jitter added and are coloured according to whether they exceed the threshold $R^*=1$, with the percentage below this threshold annotated. The point-ranges shown indicate the 25\%, 50\% and 75\% quantiles for each model.}
	\label{fig:ar1_sensitivity}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=0.6\textwidth]{bivariate.pdf}}
	\caption{\textbf{Bivariate normal example.} The distribution for $R^*$ across 1000 draws.}
	\label{fig:bivariate}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1\textwidth]{mvt_three.pdf}}
	\caption{\textbf{Multivariate normal example with 250 dimensions.} A shows $R^*$ distributions obtained for two MCMC samples (of differing numbers of draws: 400 and 10,000) from the centered parameterisation (``cp'') and one from the non-centered version (``ncp''; with 10,000 draws); B shows the rank-normalised split-$\widehat{R}$ values for all parameters from the same MCMC runs as in A; C shows variable importance versus $\widehat{R}$ for each parameter; and D shows variable importance versus bulk- and tail-ESS as calculated by \cite{vehtari2019rank}. In A, 1000 $R^*$ draws by Algorithm \ref{alg:R_star_uncertainty} are shown for each MCMC run. In plots C and D, horizontal jitter was added to the points and a loess fit line with standard errors overlaid.}
	\label{fig:mvt}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{cauchy.pdf}}
	\caption{\textbf{Cauchy example.} Rows show convergence results for MCMC runs with 1000 (top) and 10,000 (bottom; obtained by thinning iterations by a factor of 3) post-warm-up iterations (each with half iterations discarded as warm-up) for each of 4 chains. Columns show the $R^*$ distributions (left) and rank-normalised split-$\widehat{R}$ values across all parameters (right). Shadings indicate different model paramerisations as indicated in legend.}
	\label{fig:cauchy}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{cauchy_convergence.pdf}}
	\caption{\textbf{Measuring convergence for the Cauchy model.} A shows a measure of convergence, the mean quantile $R^2$, that requires knowing the target distribution; B shows the maximum value of split-$\widehat{R}$ across each of the 50 dimensions of the target; C shows the multivariate split-$\widehat{R}$ value; D shows the value of split-$R^*$ as calculated by Algorithm \ref{alg:R_star}; and E and F show tail- and bulk-ESS. Dashed lines indicate recommended thresholds for each convergence statistic.}
	\label{fig:cauchy_convergence}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{eight_schools.pdf}}
	\caption{\textbf{Eight schools example: $R^*$ distributions.} A shows draws from the $R^*$ distribution when splitting chains in two (resulting in 8 chains); B shows the same but using the 4 original chains; C shows rank-normalised $\widehat{R}$ for original 4 chains versus those for the 8 chains case for all ten parameters defined by the centered model -- in this case, we plot horizontal and vertical dashed lines to illustrate the $\widehat{R}=1.01$ cutoff and a $y=x$ line. The legend inset in panel C provides a key for all panels. The MCMC samples comprised 2000 draws in all cases, with 1000 used as post-warm-up iterations. In panels A and B, the plots show 1000 $R^*$ draws using Algorithm \ref{alg:R_star_uncertainty} for each parameterisation.}
	\label{fig:eight_schools}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{eight_schools_r_star_quantiles.pdf}}
	\caption{\textbf{Eight schools example: quantile $R^*$ plots.} Column A shows plots for $\mu$; Column B for $\tau$. In each column, we show the path of the four individual chains above and 40 $R^*$ draws obtained using Algorithm \ref{alg:R_star_uncertainty} for each parameter quantile below. Above the top row, we show the marginal distribution of each chain estimated via kernel density estimation using Gaussian kernels. Note that, in the bottom plots, jitter has been added to the data points.}
	\label{fig:eight_schools_r_star_quantiles}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1\textwidth]{mvt_wide_both.pdf}}
	\caption{\textbf{Wide data examples.} A shows the $R^*$ distribution for the 250-dimensional example in \S\ref{sec:wide} with 250 post-warm-up iterations per chain from Stan's NUTS algorithm across both model parameterisations; B shows $R^*$ distribution for the 10,000-dimensional example with 400 and 1000 MCMC iterations per chain (although the first half of these were discarded as warm-up). In all cases, 1000 draws of $R^*$ are plotted as generated by Algorithm \ref{alg:R_star_uncertainty} for a single MCMC run composed of 4 chains.}
	\label{fig:mvt_wide_both}
\end{figure}

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1\textwidth]{wide_both_diagnostics.pdf}}
	\caption{\textbf{Wide data 250-dimensional example: established diagnostics.} The top row shows the results for the centered parameterisation; the bottom row for the non-centered. Column A shows split-$\widehat{R}$; columns B and C show the bulk- and tail-ESS; in each case the statistics are displayed for all model parameters and were calculated using 250 post-warm-up draws from Stan's NUTS algorithm. Note, that it is possible for the ESS to exceed the actual sample size if there is negative autocorrelation in the Markov chains' values. In both cases, the results correspond to a single MCMC run composed of 4 chains.}
	\label{fig:wide_both_diagnostics}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{trends_all_dim.pdf}}
	\caption{\textbf{Univariate trends example.} Column A represents results for $R^*$ calculated using Algorithm \ref{alg:R_star} on the 4 chains; column B  shows the same calculation after each chain is split into two halves. The rows present the differing sample sizes. The horizontal axis measures (half) the change in mean across the whole sample: so a value ``1'' indicates the mean increases by 2 units from the start to end of sampling. At each parameter set, 10 replicates were run and jitter was added to the points.}
	\label{fig:trends_all_dim}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{trends_one_dim.pdf}}
	\caption{\textbf{Multivariate trends example with split-$R^*$.} The columns present different dimensionalities of the target distribution; the rows present different sample sizes. The horizontal axis measures (half) the change in mean across the single dimension that had a trend added to it: a value ``1'' indicates its mean increases by 2 units from the start to end of sampling; all other dimensions (if dimensions exceeded 1) had stationary distributions. At each parameter set, 10 replicates were run and jitter was added to the points.}
	\label{fig:trends_one_dim}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{trends_joint_distribution.pdf}}
	\caption{\textbf{Bivariate normal with trending correlation example.} Column A shows the results for split-$\widehat{R}$; column B for $R^*$; the rows present the differing sample sizes. The horizontal axis measures (half) the change in correlation across the whole sample: so a value ``0.5'' indicates the correlation increases by 1 unit (from -0.5 to 0.5) from the start to end of sampling. At each parameter set, 10 replicates were run and jitter was added to the points.}
	\label{fig:trends_joint_distribution}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{trends_ar1.pdf}}
	\caption{\textbf{Non-stationary distribution: AR(1) example.} The horizontal axis measures the autocorrelation of the AR(1) processes; the vertical axis shows the value of $R^*$ calculated on chains split in half; each panel shows a different number of iterations. At each parameter set, 10 replicates were run and jitter was added to the points. The black line shows $R^*=1$.}
	\label{fig:trends_ar1}
\end{figure}

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{trends_ar1_transposed.pdf}}
	\caption{\textbf{Non-stationary distribution: AR(1) example alternative view.} The horizontal axis measures the number of iterations; the vertical axis shows the value of $R^*$ calculated on chains split in half; each panel shows a different value of autocorrelation. At each parameter set, 10 replicates were run and jitter was added to the points. The black dots show the median $R^*$ values at each parameter set; upper and lower whiskers show 2.5\% and 97.5\% quantiles. The black horizontal line shows $R^*=1$. }
	\label{fig:trends_ar1_transposed}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{ovarian.pdf}}
	\caption{\textbf{Ovarian example.} In all plots, we show the results from two model runs: one with 4 chains run with 800 thinned post-warm-up iterations (9000 total iterations with 1000 discarded as warm-up; 8000 post-warm-up iterations thinned by a factor of 10); another with 16 chains run with 1000 post-warm-up iterations each (500 warm-up iterations discarded and no thinning). In A, $R^*$ distributions (with 1000 draws each using Algorithm \ref{alg:R_star_uncertainty}) are shown; B shows rank-normalised split-$\widehat{R}$ values across all parameters; C shows bulk-ESS across all parameters; and D shows tail-ESS across parameters.}
	\label{fig:ovarian}
\end{figure}

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{prostate.pdf}}
	\caption{\textbf{Prostate example.} In all plots, we show the results from two model runs: one with 4 chains run with 800 thinned post-warm-up iterations (9000 total iterations with 1000 discarded as warm-up; 8000 post-warm-up iterations thinned by a factor of 10); another with 16 chains run with 1000 post-warm-up iterations each (500 warm-up iterations discarded and no thinning). In A, $R^*$ distributions (with 1000 draws each using Algorithm \ref{alg:R_star_uncertainty}) are shown -- for the long run, these were calculated after thinning the parameters by a factor of 5; B shows rank-normalised split-$\widehat{R}$ values across all parameters; C shows bulk-ESS across all parameters; and D shows tail-ESS across parameters.}
	\label{fig:prostate}
\end{figure}
